---
title: "P120_EDA"
format: html
editor: visual
---

#### Overall data set description

Sampling for the P120 Nursery Area Juvenile Survey began in 1971. Major gear changes and standardization occured in 1978 and 1989. Sampling has been consistent in May and June (mid two weeks) throughout the sampling period. July sampled was dropped in 1998 and then reinstituded in 2004 to produce a better index for spotted seatrout. In 2008, additional habitat fields were added. There are 104 core stations sampled each year (can't be discontinued) for juvenile abundance indices. Stations sampled in addition to the 104 are selected for various purposes. Some seagrass beds along OBX are sampled May through September. The tows are one minute in duration to span 75 yards. The historical time varied (5 minutes for 20 foot net). From what I can tell, it is just one tow at each station.

Beginning in 2008 additional data parameters for Habitat were recorded: Distance to Shoreline, Shoreline/ Shoreline Structure Type, Percent Shoreline Hardened, Land Use, and Percent Cover. If distances from shore are greater than a 550 yd radius, then code as 9s. The habitat data was piloted in 2008 and additional changes were made for 2009. Bottom composition, qualitative sediment size, secchi depth, water level, water depth, water temperatures (surface and bottom), salinities (surface and bottom), and dissolved oxygen (surface and bottom) should be collected each time a station is sampled. Grass density is characterized.

All species are identified and counted. Economically important species are counted and a random subsample of 30-60 individuals of each size group are measured (mm). Blue crabs less then 20 mm are not sexed and crabs should be measured on their backs. Blue crabs are subsampled if there are more than 30 individuals that are less than 20 mm. Larger blue crabs, ones that that are greater or equal to 20 mm are sexed and all of these are measured. For other fish, if there are two different size groups and they are subsampled separately. If there are two different size groups, SpStatus will reflect this. SpStatus= 1 is for the larger size group and SpStatus= 2 is for the smaller size group (SpStatus= 0 means fish are measured as single unit). Each column is a separate species sample from a tow. To get every tow you can group by Date and Location and get an ID for the tow (id= row_number()) in order to have a row for every haul. When the Colnum is the same for a date, location, and species, there was subsampling of the species to obtain a length.

Based on the dataset description, I am going to filter P120 to start after 1989. Also, whenever you use the dataset, make sure to remove the leading 0 from location (from 2010 to 2019 dataset).

### EDA

```{r}
##Load packages and functions 
packages <- c("ggplot2", "tidyverse", "lubridate", "sf", "sp", "dplyr", "rnaturalearth", "readr", "readxl", "spatialEco", "geosphere")

invisible(lapply(packages, library, character.only= TRUE))

standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12))

world <- ne_countries(scale = "medium", returnclass = "sf")

##Load and edit data set
setwd("~/Desktop")
P120_bioledt <- read_csv("p120_biol_new.csv")

coords_captured <- function(species){
ggplot(data = world) + geom_sf() + geom_point(data = subset(p915_CPUEold, p915_CPUEold$Species %in% species & as.numeric(p915_CPUEold$Colnum > 0)), aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") + ggtitle(species)
}
```

#### General

```{r}
#Spatial orientation
##All stations
ggplot(data = world) + geom_sf() + geom_point(data = P120_bioledt, aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #CPUE

##Core stations
ggplot(data = world) + geom_sf() + geom_point(data = P120_bioledt, aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #CPUE

###Cores are 0, 1, 2, and 3: 1 and 2 are the 104 core stations sampled every year 
length(unique(P120_bioledt$Station)) #104 stations
core <- P120_bioledt %>% filter(Core == 1| Core == 2)
length(unique(core$Station)) #still 104 stations
notcore <- P120_bioledt %>% filter(Core == 3)

ggplot(data = world) + geom_sf() + geom_point(data = subset(P120_bioledt, Core==3), aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #CPUE
#core 0 looks the same as 1 or 2, core 3 has only 35 stations 

ggplot(data = world) + geom_sf() + geom_point(data = subset(P120_bioledt, Core==0), aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #CPUE

nrow(P120_bioledt %>% filter(Core== 0) %>% group_by(Date,Location) %>% mutate(id=cur_group_id()) %>% distinct(id)) #2304 samples
P120_bioledt %>% filter(Core== 0) %>% select(Month) %>% distinct(Month) #all months
nrow(P120_bioledt %>% filter(Core== 1) %>% group_by(Date,Location) %>% mutate(id=cur_group_id()) %>% distinct(id)) #7725 samples
P120_bioledt %>% filter(Core== 1) %>% select(Month) %>% distinct(Month) #5, 6 
nrow(P120_bioledt %>% filter(Core== 2) %>% group_by(Date,Location) %>% mutate(id=cur_group_id()) %>% distinct(id)) #737 samples
P120_bioledt %>% filter(Core== 2) %>% select(Month) %>% distinct(Month) #5, 6 
nrow(P120_bioledt %>% filter(Core== 3) %>% group_by(Date,Location) %>% mutate(id=cur_group_id()) %>% distinct(id)) #854 samples
P120_bioledt %>% filter(Core== 3) %>% select(Month) %>% distinct(Month) #7

#Sampling effort: general
ggplot() + geom_histogram(data = P120_bioledt, stat="count", aes(x=Ym_date)) + standard_theme #sampling picked up in 2008, I wonder why- just added more habitat fields here
ggplot() + geom_histogram(data = P120_bioledt, stat="count", aes(x=Ym_date)) + standard_theme + facet_wrap(~Core) #more sampling in core = 1 throughout the years 
ggplot() + geom_histogram(data = P120_bioledt, stat="count", aes(x=Year)) + standard_theme #more sampling in core = 1 throughout the years 

P120_bioledt %>% group_by(Date, Lat_dd, Long_dd) %>% summarize(abundance= mean(Colnum)) %>% ggplot() + geom_point(aes(x= Date, y= abundance)) + standard_theme #sum outliers with mean abundance to correct for

ggplot(data = world) + geom_sf() + geom_point(data = subset(core, between(Year, 2008, 2020)), aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd))) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude") 

ggplot(data = world) + geom_sf() + geom_point(data = subset(core, between(Year, 1989, 2007)), aes(x = as.numeric(Long_dd), y = as.numeric(Lat_dd))) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude")

#Environmental variables 
summary(is.na(P120_bioledt)) #NAs present in Grid, Quad, Time, environmental variables, Colnum, biological variables, Speciesscientificname
summary(is.na(subset(P120_bioledt, Year > 2008))) #less NAs in environmental variables after 2008

P120_bioledt_test <- P120_bioledt %>% filter(Year > 2008) %>% drop_na(Sedsize, Btmcomp) #drops only 806 data points if drop na with Sedsize and Btmcomp after 2008
```

There are a good amount of NAs in the dataset particularly with environmental variables. We might want to start analyzing the data in 2008 (or 2009?) when more of those habitat variables were included.

EDA revealed differences in sampling with consistency until 2010 and then a large increase in sampling in 2010 that didn't drop until 2020. That being said, the same stations were sampled throughout the years. I am unsure why.

Both core 0 and core 3 seem to sample a subset of stations of core 1 and 2 (near river and OBX vs. near river). Core 0 sampling occurs July- November and core 3 sampling is in July. There are 2304 samples for core 0, 7725 for core 1, 737 for core 2, and 854 for core 3. Core 3 only has 19 stations. Core station 1 and 0 have been consistently sampled. Core station 2 and 3 have had dips in sampling.Synopsis on cores: The inclusion/exclusion of cores will depend on the question. Will need to reconsider as the analysis carries on.

#### Species specific: spatial and temporal

```{r}
coords_captured <- function(species){
ggplot(data = world) + geom_sf() + geom_point(data = subset(P120_bioledt, P120_bioledt$Speciescommonname %in% species & as.numeric(P120_bioledt$Colnum > 0)), aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") + ggtitle(species)
}


coords_captured("brown shrimp")
coords_captured("bluefish")
coords_captured("atlantic croaker")
coords_captured("southern flounder")
coords_captured("red drum") #red drum are present but not that common (573 rows)
coords_captured("black drum") #red drum are present but not that common (99 rows)
coords_captured("blue catfish")
coords_captured("gizzard shad") #blue catfish, shads, and cownose ray not around 
coords_captured("american shad")
coords_captured("cownose ray")
coords_captured("blue crab")

P120_bioledt %>% filter(Speciescommonname %in% "black drum") %>% mutate(count= n())
#Depth distribution for species
t <- P120_bioledt %>% drop_na(Depth)
min(t$Depth) #0.1 m 
max(t$Depth) #9.9 m

#Temporal: for P915 you know it's not a function of sampling since they sample year around
interest_spp <- c("northern brown shrimp", "bluefish", "atlantic croaker", "southern flounder", "red drum", "black drum", "blue catfish", "gizzard shad", "american shad", "cownose ray", "blue crab", "atlantic menhaden", "spot", "summer flounder", "shads", "american eel")
fine <- P120_bioledt %>% filter(Speciescommonname %in% interest_spp, Month == 5| Month==6, Year > 2000) #species of interest
ggplot() + geom_histogram(data = subset(fine, fine$Colnum > 0), stat="count", aes(x=factor(Month))) + standard_theme + facet_wrap(~Speciescommonname, scales= "free") + xlab("Month")

#Number of observations for species of interest
test <- fine %>% group_by(Speciescommonname) %>% drop_na(Colnum) %>% summarise(sum_colnum= sum(Colnum)) #atlantic croaker, atlantic menhaden, blue crab, bluefish, northern brown shrimp, southern flounder, spot, and summer flounder have the most data 
```

#### Exploring length

```{r}
colnames(P120_bioledt)

P120_bioledt %>% filter(Speciescommonname %in% interest_spp) %>% drop_na(Length) %>% group_by(Speciescommonname) %>% summarise(min_l= min(Length), max_l= max(Length), mean_l = mean(Length), sd_l= sd(Length))

P120_bioledt %>% filter(Speciescommonname %in% interest_spp, Month == 5|Month==6) %>% drop_na(Length) %>% group_by(Speciescommonname) %>% ggplot() + geom_histogram(aes(x= Length)) + facet_wrap(~Speciescommonname, scale= "free") + xlab("Length (in)") + ylab("Count") + standard_theme
ggsave("~/Desktop/p120_length.png", width= 25, height= 25)

P120_bioledt %>% filter(Speciescommonname %in% interest_spp,Month == 5|Month==6) %>% drop_na(Length) %>% group_by(Speciescommonname, Month) %>% ggplot() + geom_histogram(aes(x= Length)) + facet_wrap(~Speciescommonname + Month, scale= "free") + xlab("Length (in)") + ylab("Count") + standard_theme
ggsave("~/Desktop/p120_length.month.png", width= 25, height= 25)
```

#### Actually dealing with length

```{r}
#Form blue crab dataset and make sure each blue crab in the subsample has their own row
P120_bc <- P120_bioledt %>% filter(Speciescommonname %in% "blue crab") %>% select(Control1, Date, Location, Colnum, Samnum, Subnum, Linenum, Frequenc, Length, Speciescommonname) %>% drop_na(Colnum, Length)
P120_bc_edt <- P120_bc %>% group_by(Control1, Colnum, Length) %>% uncount(weights=Frequenc, .remove= FALSE) %>% mutate(Frequenc_new = 1)

#Bin length: define the breaks for the intervals
values <- 1:300
breaks <- seq(0, 300, by = 10)
assigned_values <- seq(10, 300, by = 10)
result <- cut(values, breaks, labels = assigned_values) #cut values based on cut points (breaks), labels for resulting category, default is (a, b)
result_numeric <- as.numeric(as.character(result))
df <- data.frame(Length = values, Assigned_length = result_numeric)

#Calculate proportion of juveniles and adults in each Eventname
P120_bc_prop <- P120_bc_edt %>% left_join(df, by= "Length")
P120_bc_prop$Age_class <- ifelse(P120_bc_prop$Assigned_length < 50, "J", "A") #assign juvenile if under 50 mm, 50.8 mm (2 inches) is harvestable length, blue crabs should reach this length in ~1 year when they become sexually mature
P120_bc_prop <- P120_bc_prop %>% group_by(Control1, Age_class, Colnum) %>% mutate(Count= n()) %>% distinct(Count, .keep_all= TRUE) %>% ungroup() %>% group_by(Control1, Colnum) %>% mutate(Total_sampled= sum(Count)) %>% ungroup() %>% mutate(Proportion= Count/Total_sampled) %>% select(Control1, Date, Speciescommonname, Colnum, Age_class, Count, Total_sampled, Proportion)

#Now apply proportion to Colnum 
P120_bc_prop <- P120_bc_prop %>% mutate(Number= round(Colnum*Proportion,0))
P120_bc_prop$Number[P120_bc_prop$Number==0] <- 0.5 #Will have to check over this move, want it to be sampled but 
```

#For blue crabs in P120, each

#### CPUE

```{r}
#Combining datasets
##Explore species counts 
speciescount <- P120_bioledt %>% filter(Year>=2008) %>% group_by(Speciescommonname, Date, Location) %>% drop_na(Colnum) %>% distinct(Colnum) %>% ungroup() %>% group_by(Speciescommonname) %>% summarise(sum_colnum= sum(Colnum)) %>% filter(Speciescommonname %in% c("brown shrimp", "bluefish", "atlantic croaker", "southern flounder", "red drum", "black drum", "blue catfish", "gizzard shad", "american shad", "atlantic thread herring", "cownose ray", "blue crab"))
#Remove cownose ray, Atlantic thread herring, and black drum b/c n < 100, gizzard shad is only high due to one outlier, remove 
select_spp <- c("brown shrimp", "atlantic croaker", "southern flounder", "spot", "blue crab", "atlantic menhaden")

##Form a haul dataset where every row is one haul 
summary(is.na(P120_bioledt$Sciname)) #28 NAs for Sciname
P120_bioledt <- P120_bioledt %>% drop_na(Sciname)
P120_bioledt$Speciescommonname[P120_bioledt$Speciescommonname == "northern brown shrimp"] <- "brown shrimp"
length(unique(P120_bioledt$Control1)) #full dataset has more unique control1 than haul, 9641 vs. 9330, doesn't have to do with just selecting cores 
haul <- P120_bioledt %>% group_by(Date, Location) %>% mutate(ID=row_number()) %>% filter(ID==1) %>% as.data.frame() #row every haul (9310 variables whereas Lela had 8099), for all species
length(unique(haul$Control1))
length(unique(P120_bioledt$Control1))
#Let's look at which values are different: there are ~20 not in main cores, 311 values that are different--> need to ask about this 
diff_values <- setdiff(P120_bioledt$Control1, haul$Control1)

##Create dummy column for species of focus
haul$Species=NA#create a dummy column first
haul$Species=rep_along(haul$Species,c("brown shrimp", "atlantic croaker", "southern flounder", "spot", "blue crab", "atlantic menhaden", "juv blue crab", "adult blue crab")) #this repeats these species common names for the entirety of the length of the haul df
haul_df <- haul %>% dplyr::select(Control1, Program, Year, Month, Day, Date, Location, Nbrrec3, Station, Time, Duration, Sedsize, Depth, Stemp, Btemp, Ssal, Bsal, Sdo, Bdo, Secchi, Latitude, Longitude, Core, Species)
haul_df_all <- haul_df %>% tidyr::expand(Species,Control1) ##using expand to make all species associated with each collection number, is 9x (# of species) larger so it worked
haul=haul_df %>%select(-Species) #drop dummy species common name from hauldf

#now can merge haul.all and haul to create a df with a row for species associated with each collection number
haul_final=left_join(haul_df_all,haul,by="Control1")%>%as.data.frame()

sel_spp_120 <- P120_bioledt %>% filter(Speciescommonname %in% select_spp) %>% select(Control1,Date,Speciescommonname,Colnum) 
P120_bc_tojoin <- P120_bc_prop %>% select(Control1, Date, -Speciescommonname, Age_class, Number)
P120_bc_tojoin <- P120_bc_tojoin %>% mutate(Age_class= ifelse(Age_class %in% "J", "juv blue crab", "adult blue crab")) %>% rename("Speciescommonname"= "Age_class", "Colnum"= "Number")
sel_spp_120join <- rbind(sel_spp_120, P120_bc_tojoin)
sel_spp_120joinedt <- sel_spp_120join %>% group_by(Speciescommonname,Control1)%>% mutate(id=row_number()) %>% filter(id==1) %>%ungroup()%>%as.data.frame() #should always be 1 for juv. and adult blue crab b/c have distinct rows 

sel_spp_120joinedt = sel_spp_120joinedt %>% select(-id)
haul_final <- haul_final %>% rename("Speciescommonname" = "Species")
haul_sel_spp <- left_join(haul_final, sel_spp_120joinedt) %>% as.data.frame() #df is 74,480 rows, 9310 rows for each species (8 species)
haul_sel_spp$Colnum[haul_sel_spp$Colnum %in% NA] <- 0

names_join <- P120_bioledt %>% select(Speciescommonname, Sciname) %>% distinct(Speciescommonname, .keep_all= T)
p120_CPUE <- haul_sel_spp %>% left_join(names_join) %>% mutate(Sciname = ifelse(Speciescommonname == "brown shrimp", "F. aztecus", Sciname)) #brown shrimp went to NA for Sciname
p120_CPUE$Sciname <- ifelse(p120_CPUE$Speciescommonname %in% "juv blue crab"|p120_CPUE$Speciescommonname %in% "adult blue crab", "C. sapidus", p120_CPUE$Sciname)

p120_CPUE$doy <- yday(p120_CPUE$Date)
p120_CPUE$Photoperiod <- daylength(lat= p120_CPUE$Latitude, doy= p120_CPUE$doy)
p120_CPUE$Season <- ifelse(p120_CPUE$Month==4 | p120_CPUE$Month==5 | p120_CPUE$Month==6, "Spring", ifelse(p120_CPUE$Month==9 |p120_CPUE$Month==10 | p120_CPUE$Month==11 | p120_CPUE$Month==12, "Fall", ifelse(p120_CPUE$Month==7 |p120_CPUE$Month==8, "Summer", "Winter")))
p120_CPUE$Ym_date <- format(p120_CPUE$Date, "%Y-%m")

write.csv(p120_CPUE, "~/Documents/GitHub/NCBlueCrab_Predators/Data/P120/Finalized/p120_CPUE.csv")
```
