---
title: "P915DatasetSpecifics"
format: html
editor: visual
---

#### Dataset description

Sampling for the Pamlico Sound Independent Gill Net Survey began March 1st, 2001 for Hyde and Dare counties (Neuse, New, Pamlico and Pungo River systems). Sampling in the Southern District (Cape Fear and New Rivers) began in April 2008 and 2018 for Back and Core Sound. Starting in 2008, more environmental data has been recorded (e.g: sediment size, bottom composition, depth). Sampling occurs twice a month between February 15th to December 15th (January is skipped). There are 32 core samples in each month as there are 8 areas x 2x a month x 2 samples for F-70 (main region) and then for PNWGNS river system as well. For the New and Cape Fear rivers there are 12 samples per month with 8 from the New River (2 areas x 2x a month x 2 samples- shallow and deep) and 4 from Cape Fear (1 area x 2x a month x 2 shallow samples). Shallow nets are set on average of 3 feet of water and have a lot more habitat whereas deep nets are over 6 feet of water over muddy habitats.

Neuse River is divided into four areas, Pamlico River is divided into 3 areas, and Pungo river consists of 1 area. Overall, there are two regions with Region 1 encompassing areas of eastern Pamlico Sound beside the Outer Banks from southern Roanoke Island to northern end of Portsmouth Island with Region 2 including Hyde County bays from Stumpy Point to Abel's Bay and adjacent areas of western Pamlico

##### CPUE dataset

Each line of data in the CPUE dataset represents a single gillnet sampling event. When there are two rows at the same Strata on the same day, it is with the shallow and deep datasets. The set ID is represented by the control number or Sample where Sample = COMPRESS(YEAR\|\|MONTH\|\|DAY\|\|AREA\|\|QUAD\|\|GRID\|\|MODE\|\|TIME). The original datasets requested have an error with the Sample IDs. However, the Control \# is comparable to the Sample ID just utilizing numbers over the compressed format. The Control \# correlates with the record ID in the biological data set and is different for shallow/deep water sets in the same area. However, the Control \# can't be used as when P915 began, each sample was split on data sheets between half and whole mesh sizes. The different mesh sizes received different control numbers even if they were from same sampling event. The sample variable is a way to pool across the half and whole meshes on separate data sheets from earlier years. Therefore the sample id should be used to get the correct number of sampling events by year in CPUE files. Quadrants can be deep (2) or shallow (1) sets. A full day effort can be found from combining the two rows on the same strata/day with deep and shallow sets or keeping them separate. It all depends on how you want to treat the data.

However, the Control \# is representative of an individual sampling event with a unique identifier \# so it is providing the same identifier just in numbers rather than description and will correlate to record id in the other dataset. The Quad represents Deep (2) vs Shallow sets (1); so yes two rows on the same strata/day with quad 1 and 2 will be a full days effort. If you're looking to join the CPUE and overall biological dataset, control1 and record1 only works for the sampling event/environmental data. For the biological data, the record \# represents the control ID and the species ID combined so joining by area and date might be the way to go. The n/a's may stem from the raw data pull representing all effort for the program through the entire time series whereas the CPUE files are limited to years where all sampling efforts and areas occurred simultaneously (e.g. Pamlico sound started in 2001 vs rivers in 2003). In the Cape Fear River system, comparison sets have been done so it might offset the size of the samples. This shows up as REPSIZE= 1.

As of now (05/06/23), the sample IDs for the river systems (only Neuse and Pamlico) have an error. Cara Kowalchyk is planning to send me the updated CPUE data by the end of June. Right now, I can left_join by sample ID for the old CPUE and new biological dataset excluding the Neuse and Pamlico River systems.

An update about the datasets (correspondence with Cara Kowalchyk on 08/07/23): Different Control1s can be for the same sampling event because of samples split between half and whole mesh sizes (when data began). For example, 0111796 and 0111797 were the same set of deep nets but didn't merge correctly. They should be merged under sample= 11011HYDE220. If you see a sample ID that contains the control1, you can delete it off the end and it would merge properly, maybe. If you have trouble with that, Cara offered to repull the data. I propose to group_by Speciescommonname, Area, Quad, and Strata and sum the Colnum. 

However, this isn't going to be an issue because the Control1s only vary for the early data. When I did this command, p915_CPUE %>% filter(Year >= 2001) %>% group_by(Control1, Speciescommonname) %>% summarize(count = n()) and then unique(t$count), there was only 1 value for count. This means that each species only has one row for each Control1


##### Biological dataset

The raw data (biological) doesn't have a variable for no catch (0) sets. It only has data from when a species was encountered and its corresponding environmental data. The raw data is primarily used in reports for length-frequency analysis.

For each target species caught, they are 1) measured to the nearest mm fork length or total length and 2) the condition of the individual is recorded. Target species should be individually measured regardless of the \# of individuals in the sample with the exception of spot and menhaden (they should be subsampled when numbers are high). When subsampling, it is acceptable to measure 30 individuals, weigh the subsample (subsample weight), and then count and weigh the remaining specimens. It is standard to assume that the subsample weight represents the remainder of the individuals not sampled. The collection weight is the weight of all individuals collected in that gillnet set.

It is standard to divide the weight by the number of species captured. Individuals have relatively standard length throughout P915 with the exception of red drum. The first step is to bin the length and then exclude the outliers. It is important to do this for the predators selected. The prey items should always be the same size.

### EDA

```{r}
##Load packages, functions, datasets
packages <- c("ggplot2", "tidyverse", "lubridate", "sf", "sp", "dplyr", "rnaturalearth", "readr", "readxl", "spatialEco")
invisible(lapply(packages, library, character.only= TRUE))

standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12))

world <- ne_countries(scale = "medium", returnclass = "sf")

coords_captured <- function(species){
ggplot(data = world) + geom_sf() + geom_point(data = subset(p915_CPUEold, p915_CPUEold$Species %in% species & as.numeric(p915_CPUEold$Colnum > 0)), aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") + ggtitle(species)
}

setwd("~/Documents/GitHub/NCBlueCrab_Predators")
p915_CPUE <- read.csv("Data/P915/Finalized/p915_CPUE.csv")
p915_biolnew1 <- read.csv("Data/P915/Finalized/p915_biol1new_clean.csv")
p915_biolnew1 <- p915_biolnew1 %>% select(-c(X, ...1))
colnames(p915_biolnew1) <- str_to_title(colnames(p915_biolnew1))  
p915_biolnew2 <- read.csv("Data/P915/Finalized/p915_biol2new_clean.csv")
p915_biolnew2 <- p915_biolnew2 %>% select(-c(X.1, X))
colnames(p915_biolnew2) <- str_to_title(colnames(p915_biolnew2))   
p915_biolnew2 <- type.convert(p915_biolnew2, as.is = TRUE) #converts data to appropriate type, when false it converts character vectors to factors
p915_bioln <- rbind(p915_biolnew1, p915_biolnew2) #number of rows is 146079 (75496 from biolnew1 + 70583 from biolnew2), before type.convert it cut off at year 2008 
p915_bioln <- rbind(p915_biolnew1, p915_biolnew2)
p915_bioln <- p915_bioln %>% filter(!Area %in% "miss")

t <- p915_CPUE %>% filter(Control1 %in% "111797"|Control1 %in% "111796")
```

##### General

```{r}
#Spatial orientation
ggplot(data = world) + geom_sf() + geom_point(data = p915_CPUEold, aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #CPUE

ggplot(data = world) + geom_sf() + geom_point(data = p915_bioln, aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = Colnum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + labs(x= "Longitude", y= "Latitude", colour= "# caught") #Biological

unique(p915_CPUEold$Area)
unique(p915_bioln$Area)

#Sampling effort 
ggplot() + geom_histogram(data = p915_CPUEold, stat="count", aes(x=Ym_date)) + standard_theme 
ggplot() + geom_histogram(data = p915_CPUEold, stat="count", aes(x=Ym_date)) + standard_theme + facet_wrap(~Area)

p915_CPUEold %>% group_by(Date, Latitude, Longitude) %>% summarize(abundance= mean(Colnum)) %>% ggplot() + geom_point(aes(x= Date, y= abundance)) + standard_theme

ggplot() + geom_sf(data = world) + geom_point(data = p915_CPUEOld, aes(x = Longitude, y = Latitude)) + scale_size_area() + facet_wrap(Year, nrow = 6) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black"))
ggsave("/users/sallydowd/desktop/P915.sampl.yearsjpg", width=25, height= 25, units= c("cm")) #look at by year

#Environmental variables
#Presence of NAs
summary(is.na(p915_CPUE)) #NAs present in Latitude, Longitude, Sedsize, Btmcomp, Winddir, Windspd
summary(is.na(p915_CPUE)) #NAs present in Station, Quad, Time, Duration, G1_parm1, G1_parm2, environmental variables, Colnum, biological variables

p915_CPUE_test <- p915_CPUE %>% filter(Year > 2008) %>% drop_na(Sedsize) #drops 5489 points (out of 175320) when drop_na w/ Sedsize, drops way too many w/ Btmcomp

```

EDA revealed sampling consistency throughout the years (except 2020) for all areas except MHC. Sampling in MHC (Back and Core Sounds) began in 2018. One thing that I don't understand is why there are more coordinates for the biological dataset over the CPUE dataset. As all areas (except MHC) didn't overlap until 2008, it is recommended to divide by region when analyzing metrics.

After EDA, I have decided there is a need to remove NAs for latitude and longtitude, species, and colnum for both datasets. In addition, as we are only focusing on Pamlico Sound, areas in southern region can be removed (NEWR, miss, CAPEF).

Initially: excluded Neuse River and Pamlico river due to error for binding purposes.

#####Species: s&t

```{r}
#Spatial 
#Coordinates sampled at 
p915_CPUEold %>% group_by(Species) %>% summarize(n= sum(as.numeric(Colnum)))
coords_captured("bonnethead hammerhead")
coords_captured("blue crab") 
coords_captured("blue catfish") 
coords_captured("red drum") 
coords_captured("southern kingfish") 
coords_captured("southern flounder") 
coords_captured("striped bass")

#Depth distribution for species
p915_CPUEold_depth <- p915_CPUEold %>% group_by(Species, Year, Area, Quad) %>% mutate(sum= sum(Colnum)) %>% distinct(sum, .keep_all = TRUE) 

ggplot(data = world) + geom_sf() + geom_point(data = subset(p915_CPUEold_depth, p915_CPUEold_depth$Species %in% "blue catfish" & as.numeric(p915_CPUEold_depth$Colnum > 0)), aes(x = as.numeric(Longitude), y = as.numeric(Latitude), colour = sum)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black")) + facet_wrap(~Quad)

#Temporal: for P915 you know it's not a function of sampling since they sample year around
fine <- p915_CPUE %>% filter(Speciescommonname %in% c("atlantic croaker", "black drum", "bonnethead hammerhead", "red drum", "southern flounder", "southern kingfish", "sheepshead", "striped bass"), Month == 5| Month== 6| Month== 9, Year > 2000)

ggplot() + geom_histogram(data = subset(fine, fine$Colnum > 0), stat="count", aes(x=factor(Month))) + standard_theme + facet_wrap(~Speciescommonname, scales= "free") + xlab("Month")

unique(fine$Area)

ggplot(data = world) + geom_sf() + geom_point(data= t, aes(x= Longitude, y= Latitude)) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + theme(panel.background = element_rect(fill = "white", colour = "black"))
```

##### CPUE

```{r}
#Fixing the dataset mistake 
colnames(p915_CPUE)



t <- p915_CPUE %>% select(-X, -Obs, -Sedsize, -Btmcomp, -Weather, -Windspd, -Winddir, -Sciname, -Season, -Ym_date, -Photoperiod, -Wbd, -Wbdytype) %>% filter(Year > 2008) %>% group_by(Speciescommonname, Date, Area, Quad, Strata) %>% mutate(count= n()) %>% filter(Speciescommonname %in% "red drum")
write.csv(t, "/Users/sallydowd/Desktop/example.csv")
#count=2: some rows have the same environmental data and just different Colnum, others have slightly differing ones


#Some rows are straight up duplicates: count= 4, first two rows are exact same and second two are exact same, count= 3: 

#Total number of species caught 
p915_totalnum <- p915_CPUE %>% filter(Year > 2008, Month== 6, Colnum > 0) %>% group_by(Speciescommonname) %>% summarise(sum_total= sum(Colnum)) 

t2 <- t %>% filter(Colnum > 0, Month==6, Year > 2008) %>% mutate(sum_t = sum(Colnum))
unique(p915_CPUE$Area)

#CPUE for all focal species by area by year 
p915_CPUEfoc <- p915_CPUE %>% drop_na(Latitude, Longitude, Speciescommonname, Colnum) %>% select(Year, Month, Day, Area, Quad, Depth, Speciescommonname, Colnum, Latitude, Longitude, Date, Season, Depth, Ssal, Bsal, Stemp, Sedsize, Btmcomp, Season, Ym_date) %>% filter(!Speciescommonname %in% c("blue crab", "eastern oyster", "brown shrimp", "oyster toadfish", "white shrimp"))
p915_CPUEfoc <- p915_CPUEfoc %>% mutate("Soak_time" = ifelse(Area == 'CAPEF' & Month  %in% c(04, 05, 06, 07, 08, 09), 240, ifelse(Area == 'NEWR' & Month %in% c(04, 05, 06, 07, 08, 09), 240, 720))) #add in standard soak time, max is 12 hours, soak time 4 hours for CAPEF and NEWR b/w April and September
p915_CPUEfoc$CPUE= p915_CPUEfoc$Colnum/p915_CPUEfoc$Soak_time #calculate CPUE



CPUE_graph <- function(df, species, CPUE_metricn){
df %>% filter(Species %in% species) %>% ggplot(aes(x= Year, y= CPUE_avg)) +
geom_line() + geom_point() + xlab("Year") + ylab(CPUE_metricn) + standard_theme
}


#By year: doesn't produce clear trends 
p915_CPUEfoc_yr <- p915_CPUEfoc %>% group_by(Species, Year) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)
CPUE_graph(p915_CPUEfoc_yr, "bonnethead hammerhead", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "atlantic croaker", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "black drum", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "red drum", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "sheepshead", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "southern flounder", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "southern kingfish", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "striped bass", "Average CPUE") 
CPUE_graph(p915_CPUEfoc_yr, "blue catfish", "Average CPUE") 

#By year, area: some trends 
p915_CPUEfoc_yra <- p915_CPUEfoc %>% group_by(Species, Year, Area) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)

CPUE_graph(p915_CPUEfoc_yra, "bonnethead hammerhead", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "atlantic croaker", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "black drum", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "red drum", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "sheepshead", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "southern flounder", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "southern kingfish", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "striped bass", "Average CPUE") + facet_wrap(~Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yra, "blue catfish", "Average CPUE") + facet_wrap(~Area, scales= "free")

#By season, year
p915_CPUEfoc_yrs <- p915_CPUEfoc %>% group_by(Species, Season, Year) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)
CPUE_graph(p915_CPUEfoc_yrs, "bonnethead hammerhead", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "atlantic croaker", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "black drum", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "red drum", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "sheepshead", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "southern flounder", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "southern kingfish", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "striped bass", "Average CPUE") + facet_wrap(~Season, scales= "free")
CPUE_graph(p915_CPUEfoc_yrs, "blue catfish", "Average CPUE") + facet_wrap(~Season, scales= "free")

#By season, area 
p915_CPUEfoc_yrsa <- p915_CPUEfoc %>% group_by(Species, Season, Year, Area) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)

CPUE_graph(p915_CPUEfoc_yrsa, "bonnethead hammerhead", "Average CPUE") + facet_wrap(~Season + Area, scales= "free")
ggsave("/users/sallydowd/desktop/allspecs.P915.jpg", width=25, height= 25, units= c("cm"))

CPUE_graph(p915_CPUEfoc_yrsa, "red drum", "Average CPUE") + facet_wrap(~Season + Area, scales= "free")
ggsave("/users/sallydowd/desktop/allspecs.P915.jpg", width=25, height= 25, units= c("cm"))

CPUE_graph(p915_CPUEfoc_yrsa, "blue catfish", "Average CPUE") + facet_wrap(~Season + Area, scales= "free")
ggsave("/users/sallydowd/desktop/allspecs.P915.jpg", width=25, height= 25, units= c("cm"))

#By depth, year
p915_CPUEfoc_yrd <- p915_CPUEfoc %>% group_by(Species, Year, Quad) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)
CPUE_graph(p915_CPUEfoc_yrd, "bonnethead hammerhead", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "atlantic croaker", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "black drum", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "red drum", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "sheepshead", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "southern flounder", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "southern kingfish", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "striped bass", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrd, "blue catfish", "Average CPUE") + facet_wrap(~Quad, scales= "free")

#By depth, area, year
p915_CPUEfoc_yrda <- p915_CPUEfoc %>% group_by(Species, Year, Quad, Area) %>% summarise(CPUE_avg = mean(CPUE)) %>% distinct(CPUE_avg, .keep_all= TRUE)
CPUE_graph(p915_CPUEfoc_yrda, "bonnethead hammerhead", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "atlantic croaker", "Average CPUE") + facet_wrap(~Quad, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "black drum", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "red drum", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "sheepshead", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "southern flounder", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "southern kingfish", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "striped bass", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
CPUE_graph(p915_CPUEfoc_yrda, "blue catfish", "Average CPUE") + facet_wrap(~Quad + Area, scales= "free")
```

##### Centroid shifts

```{r}
#Month, Year
p915_bioln_wt <- p915_bioln %>% group_by(Speciescommonname, Latitude, Longitude, Year, Month, Ym_date) %>% summarise(Weight= mean(Colwgt))

cent_fish_nc_m <- p915_bioln_wt %>% group_by(Year, Month, Speciescommonname) %>% 
summarise(lat = questionr::wtd.mean(Latitude, Weight, na.rm = T), lon = questionr::wtd.mean(Longitude, Weight, na.rm = T)) #wtd.mean: compute the weighted mean and variance of a vector

cent_fish_dist <- function(df, species){
df %>% dplyr::filter(Speciescommonname == species) %>% ggplot() + geom_sf(data = world) + geom_point(aes(x = lon, y = lat, colour= Year), size = 1) + coord_sf(xlim=c(-78, -75), ylim=c(34,37), expand = TRUE) + standard_theme + facet_wrap(~Month) 
}

cent_fish_dist(cent_fish_nc_m, "blue crab")
cent_fish_dist(cent_fish_nc_m, "southern flounder")
cent_fish_dist(cent_fish_nc_m, "red drum")
cent_fish_dist(cent_fish_nc_m, "black drum")
cent_fish_dist(cent_fish_nc_m, "sheepshead")
cent_fish_dist(cent_fish_nc_m, "southern kingfish")
cent_fish_dist(cent_fish_nc_m, "blue catfish")
cent_fish_dist(cent_fish_nc_m, "bonnethead")

#Year
library(parameters)
library(ggpmisc)
dfwbiomass <-  p915_bioln %>% group_by(Year, Speciescommonname) %>% mutate(annualspeciesbiomass = sum(Colwgt), biomass = Colwgt)

avyearlybiomass <- dfwbiomass %>% group_by(Year, Speciescommonname) %>% summarise(mean = mean(biomass, na.rm = TRUE), se = standard_error(na.omit(biomass)))

ggplot(avyearlybiomass, aes(x = Year, y = mean)) + facet_wrap(~Speciescommonname, scales = "free") + geom_point() + geom_line() +  geom_errorbar(aes(ymax = mean + se, ymin = mean - se)) 

##Centroid of biomass: maybe add in temperature too! 
centerofbiomass <- dfwbiomass  %>% 
  mutate(weightedLAT = (biomass/annualspeciesbiomass)*Latitude) %>% 
  mutate(weightedLON = (biomass/annualspeciesbiomass)*Longitude) %>% 
  group_by(Speciescommonname, Year) %>% 
  summarise(CENTER_LAT = sum(weightedLAT, na.rm = TRUE), CENTER_LON = sum(weightedLON, na.rm = TRUE))

centerofbiomass$CENTER_LAT <- replace(centerofbiomass$CENTER_LAT, centerofbiomass$CENTER_LAT == 0, NA)
centerofbiomass$CENTER_LON <- replace(centerofbiomass$CENTER_LON, centerofbiomass$CENTER_LON == 0, NA)
ggplot(data = centerofbiomass, aes(x = Year, y = CENTER_LAT)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Center of Biomass Changes in Latitude over Time") + facet_wrap(~Speciescommonname) + xlab("Year") + ylab("Longitude") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + standard_theme

ggplot(data = centerofbiomass, aes(x = Year, y = CENTER_LON)) + geom_point() + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05, size = 2.4) + labs(title = "Center of Biomass Changes in Longitude over Time") + facet_wrap(~Speciescommonname)  + xlab("Year") + ylab("Longitude") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + standard_theme
```

#### Length/weight divisions

```{r}
#For each species look at length ranges of individuals 
p915_bioln_sel <- p915_bioln %>% dplyr::select(Control1, Date, Month, Year, Ym_date, Location, Total.species.sobserved.in.set, Station, Quad, Repsize, Sciname, Speciescommonname, Colnum, Colwgt, Samnum, Samwgt, Subnum, Subwgt, Frequenc, Length, Weight, Maturity, Age, Sample, Latitude, Longitude) %>% filter(!Speciescommonname %in% c("blue crab", "eastern oyster", "brown shrimp", "oyster toadfish", "white shrimp", NA))

summary(is.na(p915_bioln_sel)) #2 NAs in Colnum, 19 in Quad, 343 in Colwgt, 396 in Length, a lot in weight, maturity and age 

p915_bioln_sel %>% drop_na(Length) %>% group_by(Speciescommonname) %>% summarise(mean_l = mean(Length*0.0393701), min_l = min(Length*0.0393701), max_l= max(Length*0.0393701), sd_l= sd(Length*0.0393701))

p915_bioln_sel %>% drop_na(Length) %>% group_by(Speciescommonname) %>% ggplot() + geom_histogram(aes(x= Length)) + facet_wrap(~Speciescommonname, scale= "free") + xlab("Length (mm)") + ylab("Count") + standard_theme
ggsave("~/Desktop/p915_length.png", width= 25, height= 25)

p915_bioln_sel %>% drop_na(Length) %>% filter(Month== 5|Month==6|Month==9) %>% group_by(Speciescommonname, Month) %>% ggplot() + geom_histogram(aes(x= Length)) + facet_wrap(~Speciescommonname + Month, scale= "free") + xlab("Length (mm)") + ylab("Count") + standard_theme
ggsave("~/Desktop/p915_length_month.png", width= 25, height= 25)

length(unique(P120_bioledt$Control1))
length(unique(P120_bioledt$Location))
length(unique(P120_bioledt$Station))

#Expanded length frequency for all focal species by area: remove NEUSE and PAMLI 

#Expanded weight frequency for all focal species by area: remove NEUSE and PAMLI 

#centroid shifts by size class 
```

#### Spatial seperation: maybe could do this by age class, showing thorughout the years where they are located--\> centroid? in Sarah's code

#### GLM for species total number

```{r}
summary(p915_CPUEfoc)
p915_CPUEfoc$Stemp <- as.numeric(p915_CPUEfoc$Stemp)
p915_CPUEfoc$Bsal <- as.numeric(p915_CPUEfoc$Bsal)
p915_CPUEfoc$Ssal <- as.numeric(p915_CPUEfoc$Ssal)

P915_bluecatfish <- p915_CPUEfoc %>% filter(Speciescommonname %in% "blue catfish") %>% dplyr::select(Month, Year, Depth, Colnum, Latitude, Longitude, Stemp, Bsal, Ssal)
moon_pois <- glm(Colnum ~ Month + Year + Depth + Latitude + Longitude + Stemp + Bsal + Ssal, data = P915_bluecatfish, family="poisson")
summary(moon_pois)
report(moon_pois)

P915_bluecatfish <- p915_CPUEfoc %>% filter(Speciescommonname %in% "bonnethead hammerhead") %>% dplyr::select(Month, Year, Depth, Colnum, Latitude, Longitude, Stemp, Bsal, Ssal)
moon_pois <- glm(Colnum ~ Month + Year + Depth + Latitude + Longitude + Stemp + Ssal, data = P915_bluecatfish, family="poisson")
summary(moon_pois)
report(moon_pois)

unique(p915_CPUEfoc$Speciescommonname)

??report()
library(report)

P915_bluecatfish$Ssal <- as.numeric(P915_bluecatfish$Ssal)
P915_bluecatfish$Bsal <- as.numeric(P915_bluecatfish$Bsal)
P915_bluecatfish$Stemp <- as.numeric(P915_bluecatfish$Stemp)
P915_bluecatfish$Sedsize <- as.numeric(P915_bluecatfish$Sedsize)

mod_bh <- set_engine(decision_tree(mode = "regression"), "rpart")
tflow_bh <- P915_bluecatfish %>% tidyflow(seed = 23151) %>% #begin the tidyflow
plug_split(initial_split, prop = 3/4) %>% #specifies type of splitting, keep 3/4 of data for training and rest is for testing 
plug_formula(Colnum ~ Depth + Latitude + Longitude + Ssal + Bsal + Stemp + Sedsize + Month + Year) %>% #plug formula
plug_model(mod_bh) #plug the model

vanilla_fit_bh <- fit(tflow_bh)
tree_bh <- pull_tflow_fit(vanilla_fit_bh)$fit
rpart.plot(tree_bh)

pull_tflow_bh(vanilla_fit_bh)$fit
```

#### Combining datasets

```{r}
length(unique(merged$Control1)) #Control1 in biological dataset as the tow

p915_bioln_sel <- p915_bioln %>% select(Date, Area, Quad, Depth, Sedsize:Waterlvl, Latitude, Longitude, Sciname, Colnum, Weight, Age, Sex, Maturity, Colwgt:Subwgt, Length, -Ph, -Waterlvl) 
p915_CPUE_sel <- p915_CPUE %>% select(Date, Area, Quad, Depth, Sciname, Colnum, Latitude, Longitude, Ssal:Windspd)

zero_catch <- p915_CPUE_sel %>% filter(Colnum == 0) %>% select(-Secchi) %>%
add_column(Weight = NA, Age= NA, Sex= NA, Maturity= NA, Colwgt= NA, Samnum= NA, Samwgt= NA, Subnum= NA, Subwgt= NA, Length= NA)

colnames(p915_bioln_sel)
colnames(zero_catch)

idk <- rbind(p915_bioln_sel, zero_catch) #email this to Cara 

tryit <- read_excel("~/Desktop/CPUE/CPUE_blackdrum_updates.XLS.xlsx")

p915clean_new <- read_csv("~/Documents/GitHub/NCBlueCrab_Predators/Data/P915/Finalized/p915clean_new.csv")


test <- idk %>% filter(Area %in% "DARE3", Date == "2001-10-10")
test1 <- p915_bioln_sel %>% filter(Area %in% "DARE3", Date == "2001-10-10")
test3 <- p915_CPUE %>% filter(Area %in% "DARE3", Date == "2001-10-10")
test2 <- zero_catch %>% filter(Area %in% "DARE3", Date == "2001-10-10")

  
Weight
Age
Sex
Maturity
Colwgt
Samnum
Samwgt
Subnum
Subwgt
Length 


#CPUE has secchi, biological has pH and Waterlvl (drop this for now)


colnames(p915_bioln_sel)
colnames(p915_CPUE_sel)

#CPUE has each sampling event regardless of catching species of focus or not
#bioln has length, colnum, weight, etc. for recorded individuals
#Goal: dataset with length, colnum, weight for recorded individuals and record of when individuals were not recorded


#The zeros are giving the weird results- dataset does not have repeating strata and quadrants with same species unless there is a zero, thought it was one row for each sample in CPUE dataset 
#Happens with original P915 dataset too, maybe overtthinking it? 




unique(p915_CPUE$Quad)
unique(p915_bioln$Area)
t2 <- p915_CPUE %>% select(Area) %>% distinct(Area) %>% arrange(Area)
t <- p915_bioln %>% filter(Area %in% "MHDC4")

length(unique(t$Area))
#the record \# represents the control ID and the species ID combined so joining by area and date might be the way to go.
```
