---
title: "GAMs"
format: html
editor: visual
---

#### Load packages, functions, and modify final CPUE dataset

```{r}
#Load packages and functions 
packages <- c("ggplot2", "tidyverse", "lubridate", "sf", "sp", "dplyr", "rnaturalearth", "readr", "readxl", "spatialEco", "rstatix", "viridis", "BBmisc", "corrplot", "mgcv")
invisible(lapply(packages, library, character.only= TRUE))

standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12))

#Edit dataset with standardized CPUE
CPUE_grid_avg_edt <- read.csv("~/Documents/GitHub/NCBlueCrab_Predators/Data/CPUE/CPUE_grid_avg_edt.csv")
CPUE_grid_avg_edt <- CPUE_grid_avg_edt[,-1]

##Make character environmental variables factors
CPUE_grid_avg_edt <- CPUE_grid_avg_edt %>% mutate_at(c("Sedsize_common", "ShorelineType", "ITP", "Survey"), as.factor)

##Pivot_wider: each species to have their own column 
CPUE_to_run <- CPUE_grid_avg_edt
CPUE_to_run$Speciescommonname <- gsub(" ", "_", CPUE_to_run$Speciescommonname)
df_wide <- CPUE_to_run %>% filter(Survey %in% "P915") %>% ungroup() %>% pivot_wider(names_from = "Speciescommonname", values_from = "mean_CPUE_stdzd") %>% drop_na() #this removes three rows where NAs were present, feel comfortable doing this b/c a lot of most species had at least 1 NA (some 2), could be an issue from pulling the data or 

#Create binary and count datasets 
df_binary <- df_wide %>% mutate_at(vars(15:41), ~ ifelse(. > 0, 1, 0))

df_count <- read_csv("~/Documents/GitHub/NCBlueCrab_Predators/Data/CPUE/CPUE_grid_count_avg_edt.csv")  
df_count <- df_count[,-1]
df_count$Speciescommonname <- gsub(" ", "_", df_count$Speciescommonname)
df_count_wide <- df_count %>% filter(Survey %in% "P915") %>% ungroup() %>% pivot_wider(names_from = "Speciescommonname", values_from = "avg_count") %>% drop_na()
```

#### Diagnostics

GAMs, or generalized additive models, provide more flexibility than linear models to explore ecological data. A GAM is a form of a generalized linear model with a linear predictor involving a sum of smooth functions of covariates. They have three parts: 1) a random part, which describes the variability in data by means of an exponential-family probability distribution 2) a structural part which is a linear model 3) a link function that connects the structural and random part. Transformations here are applied to the predictions of a linear model. The link is different depending on the regression used: identity for linear regression, log for poisson regression (counts), and logit for logistic regression (0 or 1).

The k value is the number of basis functions that determines how wiggly the model is (smoothing parameter of a spline). A spline is a piece wise polynomial functions that are used as a smoothing technique. A rule of thumb is that it should be between 5 and the number of years in the dataset (start with 2\*number of decades in dataset). A low k-index (\< 1) might indicate that k is too low especially if edf is close to k'. Change k from there and see if p-value increases/k-index, compare k and edf values in addition to looking at p value. Small and significant p values indicate that residuals are not randomly distributed as there are not enough basis functions.

Smoothers are added to continuous variables in GAMs which is denoted by s() to fit the data. The s operator produces the spline base expansion. Different splines can be added to the model. One example is bs= "ts" which is a thin plate spline that is a smoothing basis for predictor variables. GAM in R can only accept spline smoothers (number of knots) and not LOESS. It is important to have a good compromise between roughness and smoothness. The parameter λ specifies how much weight the roughness penalty will get.

How do we know when to drop a model? If the answers to all three of these quesitons are yes, then drop it:

1) Are the EDF for the questioned term approaching 1 (i.e., linear effect)? 
2) Is the confidence interval of the term including zero throughout its domain? (i.e., high p-value) 
3) Does the GCV (generalized cross validation) score for the entire model drop when the term is removed?

#### GAM output:
plot(model3,pages=1,se=T, residuals=T,pch=1), forms covariate plots 
Use the F-ratio test 


#### Initial models

###test for over and underdispersion for poisson vs. binomial

```{r}
#Create training and test datasets 
#Create Training and validation datasets
sample_size = floor(0.8*nrow(df_binary)) #take 80% of rows 
set.seed(777)
picked = sample(seq_len(nrow(df_binary)),size = sample_size)
df_binary_test = df_binary[-picked,]
df_binary_train = df_binary[picked,]
picked2 = sample(seq_len(nrow(df_count_wide)),size = sample_size)
df_count_test = df_count_wide[-picked,]
df_count_train = df_count_wide[picked,]

#Binomial GAM with logit link
##This is a submodel that relates presence or absence to predictor variables
df_binary_train <- df_binary %>% mutate_at(vars(15:41), as.integer)
binomial <- gam(summer_flounder ~  s(avg_depth) + s(avg_ssal) + s(avg_stemp) + s(avg_sdo) + s(SAVDist_km) + s(InletDist_km) + s(NoFishRest), family=binomial(link="logit"), data= df_binary_train)
summary(binomial)
gam.check(binomial)

##Visualize impact of predictor variables 
par(mfrow=c(2,3),mai=c(0.3,0.3,0,0),omi=c(0.35,0.35,0.1,0.1))
plot(binomial, shade=TRUE, seWithMean=TRUE, scale=0, all.terms=TRUE)

#Poisson model did not work because it is for integers
poisson <- gam(summer_flounder ~  s(avg_depth) + s(avg_ssal) + s(avg_stemp) + s(avg_sdo) + s(SAVDist_km) + s(InletDist_km) + s(NoFishRest), family=poisson(), data= df_count_wide)
summary(poisson)
gam.check(poisson)
```

```{r}
gam_diagn <- function(dep_var, k_value, df){
mod <- gam(dep_var ~ s(avg_ssal, k= k_value), data = df)
print(summary(mod))
gam.check(mod)
}

#Create Training and validation datasets
sample_size = floor(0.8*nrow(CPUE_to_runP915)) #take 80% of rows 
set.seed(777)

# randomly split data in r
picked = sample(seq_len(nrow(CPUE_to_runP915)),size = sample_size)
CPUE_to_runP915Validation=CPUE_to_runP915[-picked,]

CPUE_to_runP915_updated =CPUE_to_runP915[picked,]
colnames(CPUE_to_runP915_updated)
Bmod.beinf_test <- gamlss(bonnethead_shark ~pb(InletDist_km, control=pb.control(method="GAIC")) +  pb(red_drum, control=pb.control(method="GAIC")) + 
                                             pb(SAVDist_km, control=pb.control(method="GAIC")) + 
                                             pb(NoFishRest, control=pb.control(method="GAIC")) +
                                             pb(avg_depth, control=pb.control(method="GAIC")) +
                                             pb(avg_ssal, control=pb.control(method="GAIC")) +
                                             pb(avg_stemp, control=pb.control(method="GAIC")) +
                                             pb(avg_sdo, control=pb.control(method="GAIC")) + Sedsize_common + ShorelineType, data=na.omit(CPUE_to_runP915_updated),family=BEINF0,n.cyc=ncyc, c.crit=ccrit)
summary(Bmod.beinf_test)   

```

#### Generalized linear model

```{r}
#Perfect separation is occuring 
flounder_glm <- glm(summer_flounder ~ avg_depth + avg_ssal + avg_stemp + avg_sdo + SAVDist_km + InletDist_km + NoFishRest, data = CPUE_to_runP915_updated, family="poisson") 
summary(flounder_glm)
report(flounder_glm)

autoplot.countreg(flounder_glm)
tweedie <- gam(summer_flounder ~ s(avg_depth, bs="ts", k=5) + s(avg_ssal, bs="ts", k=5) + s(avg_stemp, bs="ts", k=5) + s(avg_sdo, bs="ts", k=5) + s(SAVDist_km, bs="ts", k=5) + s(InletDist_km, bs="ts", k=5) + s(NoFishRest, bs="ts", k=5) + s(black_drum, bs="ts", k=5) + factor(FishingAll_num) + factor(Sedsize_common), family= tw(link= "log"), data= CPUE_to_runP915_updated, method= "REML")
```

#### Tweedie

The Tweedie distribution is an example of an exponential dispersion model where there is a common mathematical structure such as a link function. This distribution contains three parameters: the mean, standard deviation, and p which can stand for continuous normal, gamma, Poisson, or inverse Gaussian distributions. While the probability density function can't be evaluated with this distribution, a special algorithm is created for density calculation.

For tw() in the mgcv package, the variance function powers are between 1 (Poisson) and 2 (gamma). This power can be used to specify a distribution or can be estimated in the model. The best smoother parameter estimation method fitting method is either maximum likelihood (ML) or restricted maximum likelihood (REML) as they avoid undersmoothing. With maximum likelihood, parameters are estimated by maximizing the likelihood function. As opposed to ML, REML assumes fixed effects are known and only estimates variance components of random effects. Since ML estimates fixed and random effects simultaneously, biased estimates of variance components of random effects occur (https://aitechtrend.com/choosing-the-right-statistical-method-maximum-likelihood-vs-reml/#google_vignette).

Here are some helpful resources on Tweedie distributions in general: https://r.qcbs.ca/workshop08/book-en/other-distributions.html https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/Tweedie.html

```{r}
#About the tweedie: 
#Common link function for Tweedie is the log, link function maps non-linear relationship to linear one
library(mgcv)

#Create Training and validation datasets
sample_size = floor(0.8*nrow(CPUE_to_runP915)) #take 80% of rows 
set.seed(777)
#Randomly split data in r
picked = sample(seq_len(nrow(df_wide)),size = sample_size)
CPUE_to_runP915Validation=df_wide[-picked,]
CPUE_to_runP915_updated =df_wide[picked,]
colnames(CPUE_to_runP915_updated)

tweedie <- gam(summer_flounder ~ s(avg_depth, bs="ts", k=5) + s(avg_ssal, bs="ts", k=5) + s(avg_stemp, bs="ts", k=5) + s(avg_sdo, bs="ts", k=5) + s(SAVDist_km, bs="ts", k=5) + s(InletDist_km, bs="ts", k=5) + s(NoFishRest, bs="ts", k=5) + s(black_drum, bs="ts", k=5) + factor(FishingAll_num) + factor(Sedsize_common), family= tw(link= "log"), data= CPUE_to_runP915_updated, method= "REML")
summary(tweedie)    

smooth_gam <- gam(red_drum ~ s(avg_depth, bs="ts", k=5) + s(avg_ssal, bs="ts", k=5) + s(avg_stemp, bs="ts", k=5) + s(avg_sdo, bs="ts", k=5) + Sedsize_common, family= gaussian(link= "identity"), data= CPUE_to_runP915_updated, method= "REML")
summary(smooth_gam) 

summary(smooth_gam)$s.table
summary(tweedie)$s.table
##check the choice of k
k.check(tweedie)
k.check(smooth_gam)
##look at residual plots 
gam.check(smooth_gam)
gam.check(tweedie)
##compare the two models 
AIC(smooth_gam,tweedie)

autoplot.countreg(tweedie)

rootogram.plot <- rootogram(tweedie, style = "hanging", plot = FALSE)
```

#### Delta models

Delta (hurdle models) are a two part model with one model for zero vs. non-zero data and another model for the positive component. Hurdle models are more appropriate than something like a Tweedie when there are differences in the processes controlling presence vs. abundance, or when greater flexibility to account for dispersion is required. As ecological data often has more zeros than is expected if the process generating their data was only from a standard probability distribution, there are multiple methods for dealing with zero inflation. One of these methods is the hurdle model where the zero and non-zero data are in one model and the non-zero data is in another.

There can be Delta-gamma, Delta-lognormal, Delta-NB1 (negative binomial), and Delta-NB2 model types. With ziplss() in gam(), a zero inflated (hurdle) Poisson location-scale model family is fit to data. One linear predictor in the model controls the probability of presence (binomial) and the other controls the mean given presence with ocunts (poisson). Within a list of formulae, the linear predictors are used. The first formulae is for the linear predictor for the Poisson parameter and the second is for the probability of presence. These hurdle models should only be used when 1) there are a large number of zero counts in the data and 2) the positive counts approximate a Poisson distribution.

Helpful resources: -https://pbs-assess.github.io/sdmTMB/articles/delta-models.html -Modeling relative abundance: https://cornelllabofornithology.github.io/ebird-best-practices/abundance.html

```{r}
prop_0 <- 100*sum(CPUE_to_runP915_updated$summer_flounder == 0)/nrow(CPUE_to_runP915_updated) #definitely zero inflated 

df_count_wide_sf <- df_count_wide %>% drop_na()
df_count_wide_sf$summer_flounder <- as.integer(df_count_wide_sf$summer_flounder)

delta <- gam(list(summer_flounder ~ s(avg_depth, bs="ts", k=5) + s(avg_ssal, bs="ts", k=5) + s(avg_stemp, bs="ts", k=5) + s(avg_sdo, bs="ts", k=5) + s(SAVDist_km, bs="ts", k=5) + s(InletDist_km, bs="ts", k=5) + s(NoFishRest, bs="ts", k=5) + s(black_drum, bs="ts", k=5), ~ s(avg_depth, bs="ts", k=5) + s(avg_ssal, bs="ts", k=5) + s(avg_stemp, bs="ts", k=5) + s(avg_sdo, bs="ts", k=5) + s(SAVDist_km, bs="ts", k=5) + s(InletDist_km, bs="ts", k=5) + s(NoFishRest, bs="ts", k=5) + s(black_drum, bs="ts", k=5)), family= ziplss(), data= df_count_train, method = "REML")
summary(delta)

df_count_test = df_count_wide[-picked,]
df_count_train = df_count_wide[picked,]


# presence probability is on the complimentary log-log scale
# we can get the inverse link function with
obs_count <- select(df_count_test, obs = summer_flounder)

inv_link <- binomial(link = "cloglog")$linkinv
# combine ziplss presence and count predictions
delta_pred <- predict(delta, df_count_test, type = "link") %>% 
  as.data.frame() %>%
  transmute(family = "Zero-inflated Poisson",
            pred = inv_link(V2) * exp(V1)) %>% 
  bind_cols(obs_count)


ticks <- c(0, 1, 10, 100, 1000)
mx <- round(max(delta_pred$obs_count))
ggplot(delta_pred) +
  aes(x = log10(obs + 1), 
      y = log10(pred + 1)) +
  geom_jitter(alpha = 0.2, height = 0) +
  # y = x line
  geom_abline(slope = 1, intercept = 0, alpha = 0.5) +
  # area where counts off by a factor of 10
  geom_area(data = tibble(x = log10(seq(0, mx - 1) + 1), 
                          y = log10(seq(0, mx - 1) / 10 + 1)),
            mapping = aes(x = x, y = y),
            fill = "red", alpha = 0.2) +
  # loess fit
  geom_smooth(method = "loess", 
              method.args = list(span = 2 / 3, degree = 1)) +
  scale_x_continuous(breaks = log10(ticks + 1), labels = ticks) +
  scale_y_continuous(breaks = log10(ticks + 1), labels = ticks) +
  labs(x = "Observed count",
       y = "Predicted count") +
  facet_wrap(~ family, nrow = 1)
```

```{r}
pred.bin <- predict(binomial, df_binary_test, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(data=NA,nrow= nrow(df_binary_test),ncol=niter)
for(i in 1:nrow(df_binary_test)){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975)) 

# to get 95% CIs
plot(df_binary_test$avg_stemp, pred.mean, type="p",mgp=c(2,0.7,0), ylim=c(0,0.31),xlab="",lwd=2,cex.lab=1.4,ylab="",boxwex=0.01,cex.axis=1.1)
points(grid.bin$year,pred.mean, type="p",cex=1.5,pch=19)
points(grid.bin$year,pred.mean, type="l")
lines(grid.bin$year, pred.95CI[1,], lty=2)
lines(grid.bin$year, pred.95CI[2,], lty=2)
```
