---
title: "Part1Inference"
format: html
editor: visual
---

This script is used to calculate AIC and DIC for various generalized additive models (GAMs) and gJam models for all predators. 

The workflow is as follows: 
1) Load in script with datasets and necessary functions
2) Run GAMs separately for each predator for model types 1-6) in part 1 of the study (see Fig. 2 in manuscript)
3) Check GAM model performance and calculate AIC values
4) Run gJam models seperately for each predator for model type 3-6) in part 1 of the study (Fig. 2) based on burn-in time found in P1P2_burnin.qmd
5) Calculate DIC values for the various gJam models. 

```{r}
source("~/Documents/GitHub/NCBlueCrab_Predators/Scripts/Load.SDM.Final.R")
```

### Generalized additive models

#### Red drum

##### All models

```{r}
reddrum_env <- gam(reddrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

reddrum_bc <- gam(reddrumP915~ s(logsmallbluecrabP120) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

reddrum_forage <- gam(reddrumP915~ s(logreddrumP915forageP915) + s(logreddrumP915forageP120) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

reddrum_env_bc <- gam(reddrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logsmallbluecrabP120), family= tw(), data=df_CPUE_length_wide_both)

reddrum_env_forage <- gam(reddrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logreddrumP915forageP915+1) + s(logreddrumP915forageP120+1), family= tw(), data=df_CPUE_length_wide_both)

reddrum_env_bc_forage <- gam(reddrumP915~s(avgdepth)+s(avgssal)+s(avgstemp) + s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logreddrumP915forageP120+1) + s(logreddrumP915forageP915+1) + s(logsmallbluecrabP120+1), family= tw(),  data=df_CPUE_length_wide_both)

AIC(reddrum_env, reddrum_bc, reddrum_forage, reddrum_env_bc, reddrum_env_forage, reddrum_env_bc_forage) %>% arrange(AIC)
```

##### Diagnostics

```{r}
summary(reddrum_env)
summary(reddrum_bc)
summary(reddrum_forage)
summary(reddrum_env_bc)
summary(reddrum_env_forage)
summary(reddrum_env_bc_forage)

#check p-value for anova across years 
anova(reddrum_env)
anova(reddrum_bc)
anova(reddrum_forage)
anova(reddrum_env_bc)
anova(reddrum_env_forage)
anova(reddrum_env_bc_forage)

gam.check(reddrum_env)
gam.check(reddrum_bc)
gam.check(reddrum_forage)
gam.check(reddrum_env_bc)
gam.check(reddrum_env_forage)
gam.check(reddrum_env_bc_forage)

vif(reddrum_env)
vif(reddrum_bc)
vif(reddrum_forage)
vif(reddrum_env_bc)
vif(reddrum_env_forage)
vif(reddrum_env_bc_forage)

cor(df_CPUE_length_wide_both %>% dplyr::select(avgdepth,avgssal, avgstemp, avgsdo, NoFishRest, reddrumP915forageP915, reddrumP915forageP120, smallbluecrabP120))

#Partial effect plots
plot(reddrum_env, shade= "true") 
plot(reddrum_bc, shade= "true") 
plot(reddrum_forage, shade= "true") 
plot(reddrum_env_bc, shade= "true") 
plot(reddrum_env_forage, shade= "true") 
plot(reddrum_env_bc_forage, shade= "true") 
```

#### Southern kingfish

##### All models

```{r}
southernkingfish_env <- gam(southernkingfishP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

southernkingfish_bc <- gam(southernkingfishP915~ s(logsmallbluecrabP120) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

southernkingfish_forage <- gam(southernkingfishP915~ s(logsouthernkingfishP915forageP915) + s(logsouthernkingfishP915forageP120) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

southernkingfish_env_bc <- gam(southernkingfishP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logsmallbluecrabP120), family= tw(), data=df_CPUE_length_wide_both)

southernkingfish_env_forage <- gam(southernkingfishP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logsouthernkingfishP915forageP915) + s(logsouthernkingfishP915forageP120), family= tw(), data=df_CPUE_length_wide_both)

southernkingfish_env_bc_forage <- gam(southernkingfishP915~s(avgdepth)+s(avgssal)+s(avgstemp) + s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(logsouthernkingfishP915forageP120) + s(logsouthernkingfishP915forageP915) + s(logsmallbluecrabP120), family= tw(),  data=df_CPUE_length_wide_both)

AIC(southernkingfish_env, southernkingfish_bc, southernkingfish_forage, southernkingfish_env_bc, southernkingfish_env_forage, southernkingfish_env_bc_forage) %>% arrange(AIC)
```

##### Diagnostics

```{r}
summary(southernkingfish_env)
summary(southernkingfish_bc)
summary(southernkingfish_forage)
summary(southernkingfish_env_bc)
summary(southernkingfish_env_forage)
summary(southernkingfish_env_bc_forage)

#check p-value for anova across years 
anova(southernkingfish_env)
anova(southernkingfish_bc)
anova(southernkingfish_forage)
anova(southernkingfish_env_bc)
anova(southernkingfish_env_forage)
anova(southernkingfish_env_bc_forage)

gam.check(southernkingfish_env)
gam.check(southernkingfish_bc)
gam.check(southernkingfish_forage)
gam.check(southernkingfish_env_bc)
gam.check(southernkingfish_env_forage)
gam.check(southernkingfish_env_bc_forage)

vif(southernkingfish_env)
vif(southernkingfish_bc)
vif(southernkingfish_forage)
vif(southernkingfish_env_bc)
vif(southernkingfish_env_forage)
vif(southernkingfish_env_bc_forage)

cor(df_CPUE_length_wide_both %>% dplyr::select(avgdepth,avgssal, avgstemp, avgsdo, NoFishRest, southernkingfishP915forageP915, southernkingfishP915forageP120, smallbluecrabP120))

#Partial effect plots
plot(southernkingfish_env, shade= "true") 
plot(southernkingfish_bc, shade= "true") 
plot(southernkingfish_forage, shade= "true") 
plot(southernkingfish_env_bc, shade= "true") 
plot(southernkingfish_env_forage, shade= "true") 
plot(southernkingfish_env_bc_forage, shade= "true") 
```

#### Black drum

##### All models

```{r}
blackdrum_env <- gam(blackdrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

blackdrum_bc <- gam(blackdrumP915~ s(log(smallbluecrabP120+1)) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

blackdrum_forage <- gam(blackdrumP915~ s(log(blackdrumP915forageP120+1)) + factor(Yearfactor), family= tw(), data=df_CPUE_length_wide_both)

blackdrum_env_bc <- gam(blackdrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(log(smallbluecrabP120+1)), family= tw(), data=df_CPUE_length_wide_both)

blackdrum_env_forage <- gam(blackdrumP915~s(avgdepth)+s(avgssal)+s(avgstemp)+s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(log(blackdrumP915forageP120+1)), family= tw(), data=df_CPUE_length_wide_both)

blackdrum_env_bc_forage <- gam(blackdrumP915~s(avgdepth)+s(avgssal)+s(avgstemp) + s(avgsdo)+s(NoFishRest)+factor(Yearfactor) + s(log(blackdrumP915forageP120+1)) + s(log(smallbluecrabP120+1)), family= tw(),  data=df_CPUE_length_wide_both)

AIC(blackdrum_env, blackdrum_bc, blackdrum_forage, blackdrum_env_bc, blackdrum_env_forage, blackdrum_env_bc_forage) %>% arrange(AIC)
```

##### Diagnostics

```{r}
summary(blackdrum_env)
summary(blackdrum_bc)
summary(blackdrum_forage)
summary(blackdrum_env_bc)
summary(blackdrum_env_forage)
summary(blackdrum_env_bc_forage)

#check p-value for anova across years 
anova(blackdrum_env) 
anova(blackdrum_bc)
anova(blackdrum_forage)
anova(blackdrum_env_bc)
anova(blackdrum_env_forage)
anova(blackdrum_env_bc_forage)

gam.check(blackdrum_env)
gam.check(blackdrum_bc)
gam.check(blackdrum_forage)
gam.check(blackdrum_env_bc)
gam.check(blackdrum_env_forage)
gam.check(blackdrum_env_bc_forage)

vif(blackdrum_env)
vif(blackdrum_bc)
vif(blackdrum_forage)
vif(blackdrum_env_bc)
vif(blackdrum_env_forage)
vif(blackdrum_env_bc_forage)

cor(df_CPUE_length_wide_both %>% dplyr::select(avgdepth,avgssal, avgstemp, avgsdo, NoFishRest, blackdrumP915forageP120, smallbluecrabP120))

#Partial effect plots
plot(blackdrum_env, shade= "true") 
plot(blackdrum_bc, shade= "true") 
plot(blackdrum_forage, shade= "true") 
plot(blackdrum_env_bc, shade= "true") 
plot(blackdrum_env_forage, shade= "true") 
plot(blackdrum_env_bc_forage, shade= "true") 
```

### gJam

```{r}
df <- df_CPUE_length_wide_both

df <- df %>% mutate(logsmallbluecrabP120= log(smallbluecrabP120+1), logreddrumP915forageP915= log(reddrumP915forageP915+1), logreddrumP915forageP120= log(reddrumP915forageP120+1), logblackdrumP915forageP120= log(blackdrumP915forageP120+1), logsouthernkingfishP915forageP915= log(southernkingfishP915forageP915+1), logsouthernkingfishP915forageP120= log(southernkingfishP915forageP120+1))
```

#### Red drum

##### Environment + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(reddrumP915, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

rd_modo_env <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
rd_modo_env$fit$DIC
```

##### Environment + forage

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(reddrumP915, logreddrumP915forageP915, logreddrumP915forageP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

rd_modo_env_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
rd_modo_env_forage$fit$DIC
```

##### Environment + forage + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(reddrumP915, logreddrumP915forageP915, logreddrumP915forageP120, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

rd_modo_env_bc_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
rd_modo_env_bc_forage$fit$DIC
```

#### Southern kingfish

##### Environment + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(southernkingfishP915, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

sk_modo_env <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
sk_modo_env$fit$DIC
```

##### Environment + forage

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(southernkingfishP915, logsouthernkingfishP915forageP915, logsouthernkingfishP915forageP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

sk_modo_env_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
sk_modo_env_forage$fit$DIC
```

##### Environment + forage + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(southernkingfishP915, logsouthernkingfishP915forageP915, logsouthernkingfishP915forageP120, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

sk_modo_env_bc_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
sk_modo_env_bc_forage$fit$DIC
```

#### Black drum

##### Environment + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(blackdrumP915, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

bd_modo_env <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
bd_modo_env$fit$DIC
```

##### Environment + forage

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(blackdrumP915, logblackdrumP915forageP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

bd_modo_env_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
bd_modo_env_forage$fit$DIC
```

##### Environment + forage + BC

```{r}
xdata <- df %>% dplyr::select(avgdepth, avgstemp, avgssal, avgsdo, NoFishRest, Yearfactor)

ydata <- df %>% dplyr::select(blackdrumP915, logblackdrumP915forageP120, logsmallbluecrabP120) 

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
formula_str <- paste("~", paste(colnames(xdata), collapse = " + "))

bd_modo_env_bc_forage <- gjam(formula= as.formula(formula_str), xdata = xdata, ydata = ydata, modelList = ml)
bd_modo_env_bc_forage$fit$DIC
```
