---
title: "P2Prediction"
format: html
editor: visual
---

This script is used to run cross-validation for gJam for Model 3a in part 2 of the study (individual prey item models).

The workflow is as follows: 1) Load in script with datasets and necessary functions 2) Run gJam separately for each predator for Model 1a, Model 2a, and Model 3a in part 2 of the study (see Fig. 2 in manuscript) with 5-fold cross-validation repeated 100 times. Even though these models includes multiple species CPUE in the response, a separate model is run for each predator. This enables us to test the model's predictive capability for predators separately. 4) Load in these data. 5) Compute pairwise comparisons 6) Collect summary metrics.

```{r}
source("~/Documents/GitHub/NCBlueCrab_Predators/Scripts/Load.SDM.Final.R")
```

#### Model 1a 
```{r}
ydata <- df_CPUE_length_wide_both %>% dplyr::select(smallatlanticcroakerP915, smallatlanticmenhadenP915, blackdrumP915, pinfishP915, reddrumP915, smallsouthernflounderP915, southernkingfishP915, smallspotP915, smallatlanticcroakerP120, smallbluecrabP120, brownshrimpP120, whiteshrimpP120, pinkshrimpP120, pinfishP120, southernflounderP120, smallspotP120, atlanticmenhadenP120)
species <- gjamTrimY(ydata, 50, OTHER = FALSE)$y %>% colnames()
ydata <- ydata[, species] %>% as.data.frame() #filter out rare species
```

##### Red drum
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_both))

reddrum_model1_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(reddrum_model1_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_both)), size = smp_size)
train <- df_CPUE_length_wide_both[train_ind, ]
test <- df_CPUE_length_wide_both[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "reddrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_reddrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "reddrumP915"]
obs_reddrum <- ydata_test[,colnames(ydata_test) %in% "reddrumP915"] #observed reddrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_reddrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "reddrumP915"]
pred_reddrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "reddrumP915"]

dat <- cbind(pred_reddrum_con, obs_reddrum, pred_reddrum_un)
dat <- as.data.frame(dat)

reddrum_model1_gjam[i,1] <- r2_general(dat$pred_reddrum_con, dat$reddrumP915)
reddrum_model1_gjam[i,2] <- r2_general(dat$pred_reddrum_un, dat$reddrumP915)
reddrum_model1_gjam[i,3] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_con)
reddrum_model1_gjam[i,4] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_un) 
print(i)
}

write.csv(reddrum_model1_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/reddrum_model1.csv")
```

##### Southern kingfish
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_both))

southernkingfish_model1_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(southernkingfish_model1_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_both)), size = smp_size)
train <- df_CPUE_length_wide_both[train_ind, ]
test <- df_CPUE_length_wide_both[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "southernkingfishP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_southernkingfish_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "southernkingfishP915"]
obs_southernkingfish <- ydata_test[,colnames(ydata_test) %in% "southernkingfishP915"] #observed southernkingfish

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_southernkingfish_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "southernkingfishP915"]
pred_southernkingfish_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "southernkingfishP915"]

dat <- cbind(pred_southernkingfish_con, obs_southernkingfish, pred_southernkingfish_un)
dat <- as.data.frame(dat)

southernkingfish_model1_gjam[i,1] <- r2_general(dat$pred_southernkingfish_con, dat$southernkingfishP915)
southernkingfish_model1_gjam[i,2] <- r2_general(dat$pred_southernkingfish_un, dat$southernkingfishP915)
southernkingfish_model1_gjam[i,3] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_con)
southernkingfish_model1_gjam[i,4] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_un) 
print(i)
}

write.csv(southernkingfish_model1_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model1.csv")
```

##### Black drum
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_both))

blackdrum_model1_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(blackdrum_model1_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_both)), size = smp_size)
train <- df_CPUE_length_wide_both[train_ind, ]
test <- df_CPUE_length_wide_both[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "blackdrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_blackdrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "blackdrumP915"]
obs_blackdrum <- ydata_test[,colnames(ydata_test) %in% "blackdrumP915"] #observed blackdrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_blackdrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "blackdrumP915"]
pred_blackdrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "blackdrumP915"]

dat <- cbind(pred_blackdrum_con, obs_blackdrum, pred_blackdrum_un)
dat <- as.data.frame(dat)

blackdrum_model1_gjam[i,1] <- r2_general(dat$pred_blackdrum_con, dat$blackdrumP915)
blackdrum_model1_gjam[i,2] <- r2_general(dat$pred_blackdrum_un, dat$blackdrumP915)
blackdrum_model1_gjam[i,3] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_con)
blackdrum_model1_gjam[i,4] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_un) 
print(i)
}

write.csv(blackdrum_model1_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model1.csv")
```

#### Model 2

```{r}
#May and June, 2001-2019
ydata <- df_CPUE_length_wide_P915 %>% dplyr::select(smallatlanticcroakerP915, smallatlanticmenhadenP915, blackdrumP915, pinfishP915, reddrumP915, smallsouthernflounderP915, southernkingfishP915, smallspotP915)
species <- gjamTrimY(ydata, 50, OTHER = FALSE)$y %>% colnames()
ydata <- ydata[, species] %>% as.data.frame() #filter out rare species
```

##### Red drum

```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_P915))

reddrum_model2_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(reddrum_model2_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_P915)), size = smp_size)
train <- df_CPUE_length_wide_P915[train_ind, ]
test <- df_CPUE_length_wide_P915[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "reddrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_reddrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "reddrumP915"]
obs_reddrum <- ydata_test[,colnames(ydata_test) %in% "reddrumP915"] #observed reddrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_reddrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "reddrumP915"]
pred_reddrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "reddrumP915"]

dat <- cbind(pred_reddrum_con, obs_reddrum, pred_reddrum_un)
dat <- as.data.frame(dat)

reddrum_model2_gjam[i,1] <- r2_general(dat$pred_reddrum_con, dat$reddrumP915)
reddrum_model2_gjam[i,2] <- r2_general(dat$pred_reddrum_un, dat$reddrumP915)
reddrum_model2_gjam[i,3] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_con)
reddrum_model2_gjam[i,4] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_un) 
print(i)
}

write.csv(reddrum_model2_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/reddrum_model2.csv")
```

##### Southern kingfish
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_P915))

southernkingfish_model2_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(southernkingfish_model2_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_P915)), size = smp_size)
train <- df_CPUE_length_wide_P915[train_ind, ]
test <- df_CPUE_length_wide_P915[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "southernkingfishP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_southernkingfish_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "southernkingfishP915"]
obs_southernkingfish <- ydata_test[,colnames(ydata_test) %in% "southernkingfishP915"] #observed southernkingfish

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_southernkingfish_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "southernkingfishP915"]
pred_southernkingfish_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "southernkingfishP915"]

dat <- cbind(pred_southernkingfish_con, obs_southernkingfish, pred_southernkingfish_un)
dat <- as.data.frame(dat)

southernkingfish_model2_gjam[i,1] <- r2_general(dat$pred_southernkingfish_con, dat$southernkingfishP915)
southernkingfish_model2_gjam[i,2] <- r2_general(dat$pred_southernkingfish_un, dat$southernkingfishP915)
southernkingfish_model2_gjam[i,3] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_con)
southernkingfish_model2_gjam[i,4] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_un) 
print(i)
}

write.csv(southernkingfish_model2_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model2.csv")

mean(southernkingfish_model2_gjam$r2_uncon)
mean(southernkingfish_model2_gjam$r2_con)
sd(southernkingfish_model2_gjam$r2_uncon)
sd(southernkingfish_model2_gjam$r2_con)

mean(southernkingfish_model2_gjam$rmse_uncon)
mean(southernkingfish_model2_gjam$rmse_con)
sd(southernkingfish_model2_gjam$rmse_uncon)
sd(southernkingfish_model2_gjam$rmse_con)
```

##### Black drum
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_P915))

blackdrum_model2_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(blackdrum_model2_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_P915)), size = smp_size)
train <- df_CPUE_length_wide_P915[train_ind, ]
test <- df_CPUE_length_wide_P915[-train_ind, ]

xnames <- c("avgdepth", "avgstemp", "avgssal", "avgsdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')

mod2 <- gjam(~ avgdepth + avgstemp + avgssal + avgsdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "blackdrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_blackdrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "blackdrumP915"]
obs_blackdrum <- ydata_test[,colnames(ydata_test) %in% "blackdrumP915"] #observed blackdrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_blackdrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "blackdrumP915"]
pred_blackdrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "blackdrumP915"]

dat <- cbind(pred_blackdrum_con, obs_blackdrum, pred_blackdrum_un)
dat <- as.data.frame(dat)

blackdrum_model2_gjam[i,1] <- r2_general(dat$pred_blackdrum_con, dat$blackdrumP915)
blackdrum_model2_gjam[i,2] <- r2_general(dat$pred_blackdrum_un, dat$blackdrumP915)
blackdrum_model2_gjam[i,3] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_con)
blackdrum_model2_gjam[i,4] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_un) 
print(i)
}

write.csv(blackdrum_model2_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model2.csv")

mean(blackdrum_model2_gjam$r2_uncon)
mean(blackdrum_model2_gjam$r2_con)
sd(blackdrum_model2_gjam$r2_uncon)
sd(blackdrum_model2_gjam$r2_con)

mean(blackdrum_model2_gjam$rmse_uncon)
mean(blackdrum_model2_gjam$rmse_con)
sd(blackdrum_model2_gjam$rmse_uncon)
sd(blackdrum_model2_gjam$rmse_con)

beep(sound= 8)
```

#### Model 3
```{r}
#May and June, 2001-2019
ydata <-df_CPUE_length_wide_indP915 %>% dplyr::select(smallatlanticcroakerP915, smallatlanticmenhadenP915, blackdrumP915, pinfishP915, reddrumP915, smallsouthernflounderP915, southernkingfishP915, smallspotP915)
species <- gjamTrimY(ydata, 50, OTHER = FALSE)$y %>% colnames() #minimum # of non-zero observations: 50
ydata <- ydata[, species] %>% as.data.frame() #filter out rare species
```

##### Red drum
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_indP915))

reddrum_model3_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(reddrum_model3_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_indP915)), size = smp_size)
train <- df_CPUE_length_wide_indP915[train_ind, ]
test <- df_CPUE_length_wide_indP915[-train_ind, ]

xnames <- c("Depth", "Stemp", "Ssal", "Sdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 10000, burnin = 400, typeNames = 'CA')

mod2 <- gjam(~ Depth + Stemp + Ssal + Sdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "reddrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_reddrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "reddrumP915"]
obs_reddrum <- ydata_test[,colnames(ydata_test) %in% "reddrumP915"] #observed reddrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_reddrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "reddrumP915"]
pred_reddrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "reddrumP915"]

dat <- cbind(pred_reddrum_con, obs_reddrum, pred_reddrum_un)
dat <- as.data.frame(dat)

reddrum_model3_gjam[i,1] <- r2_general(dat$pred_reddrum_con, dat$reddrumP915)
reddrum_model3_gjam[i,2] <- r2_general(dat$pred_reddrum_un, dat$reddrumP915)
reddrum_model3_gjam[i,3] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_con)
reddrum_model3_gjam[i,4] <- RMSE_func(actual = dat$reddrumP915, pred = dat$pred_reddrum_un) 
print(i)
}

write.csv(reddrum_model3_gjam, "C:/Users/sdowd/Desktop/reddrum_model3.csv")
```

##### Southern kingfish
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_indP915))

southernkingfish_model3_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(southernkingfish_model3_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_indP915)), size = smp_size)
train <- df_CPUE_length_wide_indP915[train_ind, ]
test <- df_CPUE_length_wide_indP915[-train_ind, ]

xnames <- c("Depth", "Stemp", "Ssal", "Sdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 10000, burnin = 400, typeNames = 'CA')

mod2 <- gjam(~ Depth + Stemp + Ssal + Sdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "southernkingfishP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_southernkingfish_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "southernkingfishP915"]
obs_southernkingfish <- ydata_test[,colnames(ydata_test) %in% "southernkingfishP915"] #observed southernkingfish

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_southernkingfish_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "southernkingfishP915"]
pred_southernkingfish_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "southernkingfishP915"]

dat <- cbind(pred_southernkingfish_con, obs_southernkingfish, pred_southernkingfish_un)
dat <- as.data.frame(dat)

southernkingfish_model3_gjam[i,1] <- r2_general(dat$pred_southernkingfish_con, dat$southernkingfishP915)
southernkingfish_model3_gjam[i,2] <- r2_general(dat$pred_southernkingfish_un, dat$southernkingfishP915)
southernkingfish_model3_gjam[i,3] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_con)
southernkingfish_model3_gjam[i,4] <- RMSE_func(actual = dat$southernkingfishP915, pred = dat$pred_southernkingfish_un) 
print(i)
}

write.csv(southernkingfish_model3_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model3.csv")
```

##### Black drum
```{r}
Bootstrap_times <- 100
smp_size <- floor(0.80* nrow(df_CPUE_length_wide_indP915))

blackdrum_model3_gjam <- data.frame(matrix(ncol=4, nrow=1))
colnames(blackdrum_model3_gjam) <- c("r2_con", "r2_uncon", "rmse_con", "rmse_uncon")

for(i in 1:Bootstrap_times){
train_ind <- sample(seq_len(nrow(df_CPUE_length_wide_indP915)), size = smp_size)
train <- df_CPUE_length_wide_indP915[train_ind, ]
test <- df_CPUE_length_wide_indP915[-train_ind, ]

xnames <- c("Depth", "Stemp", "Ssal", "Sdo", "NoFishRest")
ynames <- species
xdata_train <- train[,colnames(train) %in% xnames]
ydata_train <- train[,colnames(train) %in% ynames]
xdata_test <- test  
ydata_test <- test[,colnames(test) %in% ynames]

ml <- list(ng = 10000, burnin = 400, typeNames = 'CA')

mod2 <- gjam(~ Depth + Stemp + Ssal + Sdo + NoFishRest, xdata = xdata_train, ydata = ydata_train, modelList = ml) #4/5 of data 

ycond <- ydata_test[,!colnames(ydata_test) %in% "blackdrumP915"]
newdata1 <- list(xdata = xdata_test, ydataCond = ycond, nsim = 1000) 
p6 <- gjamPredict(mod2, newdata= newdata1) #1/5 of data 
pred_blackdrum_con <- p6$sdList$yMu[,colnames(p6$sdList$yMu) %in% "blackdrumP915"]
obs_blackdrum <- ydata_test[,colnames(ydata_test) %in% "blackdrumP915"] #observed blackdrum

#unconditional prediction out of sample
newdata <- list(xdata = xdata_test, nsim = 1000)
p7 <- gjamPredict(mod2, newdata = newdata)
pred_blackdrum_un <- p7$sdList$yMu[,colnames(p7$sdList$yMu) %in% "blackdrumP915"]
pred_blackdrum_un2 <- mod2$prediction$ypredMu[,colnames(mod2$prediction$ypredMu) %in% "blackdrumP915"]

dat <- cbind(pred_blackdrum_con, obs_blackdrum, pred_blackdrum_un)
dat <- as.data.frame(dat)

blackdrum_model3_gjam[i,1] <- r2_general(dat$pred_blackdrum_con, dat$blackdrumP915)
blackdrum_model3_gjam[i,2] <- r2_general(dat$pred_blackdrum_un, dat$blackdrumP915)
blackdrum_model3_gjam[i,3] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_con)
blackdrum_model3_gjam[i,4] <- RMSE_func(actual = dat$blackdrumP915, pred = dat$pred_blackdrum_un) 
print(i)
}

write.csv(blackdrum_model3_gjam, "~/Desktop/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model3.csv")
```

#### Load in data
```{r}
#Red drum 
reddrum_model1 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/reddrum_model1.csv")
reddrum_model2 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/reddrum_model2.csv")
reddrum_model3 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/reddrum_model3.csv")
```

```{r}
#Southern kingfish
southern_kingfish_model1 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model1.csv")
southern_kingfish_model2 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model2.csv")
southern_kingfish_model3 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/southernkingfish_model3.csv")
```

```{r}
#Black drum
black_drum_model1 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model1.csv")
black_drum_model2 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model2.csv")
black_drum_model3 <- read.csv("/Users/sallydowd/Google Drive/My Drive/Research/Ch1Data/Final_results/gJam/Prediction2/blackdrum_model3.csv")
```

#### Pairwise comparisons
```{r}
### Red drum ###
#Model to model: conditional
wilcox.test(reddrum_model1$r2_con, reddrum_model2$r2_con, paired = FALSE)
wilcox.test(reddrum_model1$r2_con, reddrum_model3$r2_con, paired = FALSE)
wilcox.test(reddrum_model2$r2_con, reddrum_model3$r2_con, paired = FALSE)
wilcox.test(reddrum_model1$rmse_con, reddrum_model2$rmse_con, paired = FALSE)
wilcox.test(reddrum_model1$rmse_con, reddrum_model3$rmse_con, paired = FALSE)
wilcox.test(reddrum_model2$rmse_con, reddrum_model3$rmse_con, paired = FALSE)

#Model to model: unconditional
wilcox.test(reddrum_model1$r2_uncon, reddrum_model2$r2_uncon, paired = FALSE)
wilcox.test(reddrum_model1$r2_uncon, reddrum_model3$r2_uncon, paired = FALSE)
wilcox.test(reddrum_model2$r2_uncon, reddrum_model3$r2_uncon, paired = FALSE)
wilcox.test(reddrum_model1$rmse_uncon, reddrum_model2$rmse_uncon, paired = FALSE)
wilcox.test(reddrum_model1$rmse_uncon, reddrum_model3$rmse_uncon, paired = FALSE)
wilcox.test(reddrum_model2$rmse_uncon, reddrum_model3$rmse_uncon, paired = FALSE)

#Within model
wilcox.test(reddrum_model1$r2_con, reddrum_model1$r2_uncon, paired = TRUE)
wilcox.test(reddrum_model2$r2_con, reddrum_model2$r2_uncon, paired = TRUE)
wilcox.test(reddrum_model3$r2_con, reddrum_model3$r2_uncon, paired = TRUE)
```

```{r}
### Southern kingfish ###
#Model to model: conditional
wilcox.test(southern_kingfish_model1$r2_con, southern_kingfish_model2$r2_con, paired = FALSE)
wilcox.test(southern_kingfish_model1$r2_con, southern_kingfish_model3$r2_con, paired = FALSE)
wilcox.test(southern_kingfish_model2$r2_con, southern_kingfish_model3$r2_con, paired = FALSE)
wilcox.test(southern_kingfish_model1$rmse_con, southern_kingfish_model2$rmse_con, paired = FALSE)
wilcox.test(southern_kingfish_model1$rmse_con, southern_kingfish_model3$rmse_con, paired = FALSE)
wilcox.test(southern_kingfish_model2$rmse_con, southern_kingfish_model3$rmse_con, paired = FALSE)

#Model to model: unconditional
wilcox.test(southern_kingfish_model1$r2_uncon, southern_kingfish_model2$r2_uncon, paired = FALSE)
wilcox.test(southern_kingfish_model1$r2_uncon, southern_kingfish_model3$r2_uncon, paired = FALSE)
wilcox.test(southern_kingfish_model2$r2_uncon, southern_kingfish_model3$r2_uncon, paired = FALSE)
wilcox.test(southern_kingfish_model1$rmse_uncon, southern_kingfish_model2$rmse_uncon, paired = FALSE)
wilcox.test(southern_kingfish_model1$rmse_uncon, southern_kingfish_model3$rmse_uncon, paired = FALSE)
wilcox.test(southern_kingfish_model2$rmse_uncon, southern_kingfish_model3$rmse_uncon, paired = FALSE)

#Within model
wilcox.test(southern_kingfish_model1$r2_con, southern_kingfish_model1$r2_uncon, paired = TRUE)
wilcox.test(southern_kingfish_model2$r2_con, southern_kingfish_model2$r2_uncon, paired = TRUE)
wilcox.test(southern_kingfish_model3$r2_con, southern_kingfish_model3$r2_uncon, paired = TRUE)
```

```{r}
### Black drum ###
#Model to model: conditional
wilcox.test(black_drum_model1$r2_con, black_drum_model2$r2_con, paired = FALSE)
wilcox.test(black_drum_model1$r2_con, black_drum_model3$r2_con, paired = FALSE)
wilcox.test(black_drum_model2$r2_con, black_drum_model3$r2_con, paired = FALSE)
wilcox.test(black_drum_model1$rmse_con, black_drum_model2$rmse_con, paired = FALSE)
wilcox.test(black_drum_model1$rmse_con, black_drum_model3$rmse_con, paired = FALSE)
wilcox.test(black_drum_model2$rmse_con, black_drum_model3$rmse_con, paired = FALSE)

#Model to model: unconditional
wilcox.test(black_drum_model1$r2_uncon, black_drum_model2$r2_uncon, paired = FALSE)
wilcox.test(black_drum_model1$r2_uncon, black_drum_model3$r2_uncon, paired = FALSE)
wilcox.test(black_drum_model2$r2_uncon, black_drum_model3$r2_uncon, paired = FALSE)
wilcox.test(black_drum_model1$rmse_uncon, black_drum_model2$rmse_uncon, paired = FALSE)
wilcox.test(black_drum_model1$rmse_uncon, black_drum_model3$rmse_uncon, paired = FALSE)
wilcox.test(black_drum_model2$rmse_uncon, black_drum_model3$rmse_uncon, paired = FALSE)

#Within model
wilcox.test(black_drum_model1$r2_con, black_drum_model1$r2_uncon, paired = TRUE)
wilcox.test(black_drum_model2$r2_con, black_drum_model2$r2_uncon, paired = TRUE)
wilcox.test(black_drum_model3$r2_con, black_drum_model3$r2_uncon, paired = TRUE)
```

#### Summary metrics
```{r}
### Red drum ###
mean(reddrum_model1$r2_uncon)
sd(reddrum_model1$r2_uncon)
mean(reddrum_model1$rmse_uncon)
sd(reddrum_model1$rmse_uncon)
mean(reddrum_model1$r2_con)
sd(reddrum_model1$r2_con)
mean(reddrum_model1$rmse_con)
sd(reddrum_model1$rmse_con)

mean(reddrum_model2$r2_uncon)
sd(reddrum_model2$r2_uncon)
mean(reddrum_model2$rmse_uncon)
sd(reddrum_model2$rmse_uncon)
mean(reddrum_model2$r2_con)
sd(reddrum_model2$r2_con)
mean(reddrum_model2$rmse_con)
sd(reddrum_model2$rmse_con)

mean(reddrum_model3$r2_uncon)
sd(reddrum_model3$r2_uncon)
mean(reddrum_model3$rmse_uncon)
sd(reddrum_model3$rmse_uncon)
mean(reddrum_model3$r2_con)
sd(reddrum_model3$r2_con)
mean(reddrum_model3$rmse_con)
sd(reddrum_model3$rmse_con)
```

```{r}
### Southern kingfish ###
mean(black_drum_model1$r2_uncon)
sd(black_drum_model1$r2_uncon)
mean(black_drum_model1$rmse_uncon)
sd(black_drum_model1$rmse_uncon)
mean(black_drum_model1$r2_con)
sd(black_drum_model1$r2_con)
mean(black_drum_model1$rmse_con)
sd(black_drum_model1$rmse_con)

mean(black_drum_model2$r2_uncon)
sd(black_drum_model2$r2_uncon)
mean(black_drum_model2$rmse_uncon)
sd(black_drum_model2$rmse_uncon)
mean(black_drum_model2$r2_con)
sd(black_drum_model2$r2_con)
mean(black_drum_model2$rmse_con)
sd(black_drum_model2$rmse_con)

mean(black_drum_model3$r2_uncon)
sd(black_drum_model3$r2_uncon)
mean(black_drum_model3$rmse_uncon)
sd(black_drum_model3$rmse_uncon)
mean(black_drum_model3$r2_con)
sd(black_drum_model3$r2_con)
mean(black_drum_model3$rmse_con)
sd(black_drum_model3$rmse_con)
```
